---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

### Technical part, library load etc. 

```{r,echo=TRUE,results='hide', message=FALSE, warning=FALSE}
library(utils)
library(dplyr)
```

### Loading and preprocessing the data


```{r,echo=TRUE,results='hide', message=FALSE, warning=FALSE}
data_source <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
data_dest <- paste(getwd(), "/activity.zip", sep = "")
download.file(data_source, data_dest)
unzip(data_dest)
data_read <- paste(getwd(), "/activity.csv", sep = "")

base_data <- read.csv(data_read, header = TRUE, sep = ",")

```
Lets check if data were loaded properly:

```{r}
head(base_data)
```
And now remove "NA" rows:

```{r}
base_data_nona <- na.omit(base_data)
```
so finally data looks like:

```{r}
head(base_data_nona)
```

### What is mean total number of steps taken per day?

```{r}
total_per_day <- base_data_nona %>% group_by(date) %>% summarise(sum = sum(steps))
total_per_day
```

Now lets look and examine histogram of our total data:

```{r}
hist(total_per_day$sum, 
     main="Number of steps per day", 
     xlab = "Total number of steps per day", 
     ylab = "Frequency", 
     col="darkgreen", 
     freq = TRUE)

```

and following that mean and median of total number of steps taken each day:

```{r}
mean(total_per_day$sum)
median(total_per_day$sum)
```

### What is the average daily activity pattern?

```{r}
daily_pattern <- base_data_nona %>% group_by(interval) %>% summarise(mean = mean(steps))
```

```{r}
daily_pattern
```

Now, lets see how does it look on plot:

```{r}
plot(daily_pattern$interval, daily_pattern$mean, type = "l", main = "Avarage number of steps across a day in 5-mins intervals", xlab = "Minutes", ylab = "Avg number of steps", col="darkgreen")

```

The 5-mins interval, during which on average ther biggest number of steps is taken is as following:

```{r}
daily_pattern[daily_pattern$mean == max(daily_pattern$mean),]

```

### Imputing missing values
Lets start witch checking how many missing values do we have in our data set:

```{r}
summary(base_data)
```

As we see, we have 2304 rows with missing value on steps variable. 

Lets replace NAs with mean for day, when they occurred:

```{r}
total_per_day <- base_data %>% group_by(date) %>% summarise(sum = sum(steps))
```

### Are there differences in activity patterns between weekdays and weekends?
